---
title: How I check the data
author: "Marie Vaugoyeau"
date: '2018-02-22'
slug: how-i-check-the-data
categories:
- stat
tags:
- stat
output:
  blogdown::html_page:
    toc: true
featuredpath: "/images/featured/"
featured: "ValidationDonnees.png"
featuredalt: "How I check the data"
linktitle: "how-i-check-the-data"
images: 
  - /img/featured/ValidationDonnees.png  
description : "Check data to find misprints"
---

Actually I analyse data from a thesis which measure urbanisation's influence on physicochemical characteristics. The PhD student sample water once by month during 2017 at seven spots around a lake.  
I think it is a nice occasion to explain how I check the data when I received them.  
  
# There are outliers?
## First step: Outliers and Boxplot  
Before to start anything, I test the data to find outliers (observation that has a relatively large or small value compared to the majority of observations [Zuur et al. (2010)](http://onlinelibrary.wiley.com/doi/10.1111/j.2041-210X.2009.00001.x/abstract)).  
```{r setup, include=FALSE}
X<-read.table("PhysicochimieJanvDec2017aVerifier.csv", header = TRUE, sep = ";")
Y<-read.table("PhysicochimieJanvDec2017.csv", header = TRUE, sep = ";")
```
I start by draw boxplots by month (one by characteristic).  
```{r boxplot, echo=TRUE, message=FALSE, warning=FALSE, include=TRUE}
library(ggplot2)
for (i in 3:6)
{
bp<-ggplot(X) + geom_boxplot(aes(x = as.factor(Date), y = X[,i])) + geom_jitter(aes(x = as.factor(Date), y = X[,i]), col = X$Site, alpha = 0.2) + theme_minimal() + xlab("Date") + ylab(variable.names(X)[i])
print(bp)
}
```
Except for NO~3~^-^ at 128 days after the start of the year, there does not seem to have any outlier.  
  
## Second step: Outliers, means and standard error
In the data used, measurements were did three times each, so I calculated means and standard error with plotrix library.  
```{r mean, echo=TRUE, message=FALSE, warning=FALSE, include=TRUE}
library(plotrix)

X$rep<-paste(as.factor(X$Site), as.factor(X$Date))
# I create a variable to work on the repetition of measurements

# I create also the new database Z to have means and standard error according to spot and date
Site<-rep(1:7, each = 12)
Date<-rep(c(128,159,189,220,251,294,312,342,35,67,8,98), 7)
Z<-cbind(Site,Date)
CN<-colnames(Z)

for (i in 3:6)
{
M<-tapply(X[,i], X$rep, mean, na.rm=TRUE)
S<-tapply(X[,i], X$rep, std.error, na.rm=TRUE)
Z<-cbind(Z, M, S)
CN<-c(CN, paste(variable.names(X)[i],"Mean"),paste(variable.names(X)[i],"Se"))
}

colnames(Z)<-CN

write.table(Z, row.names=FALSE, "Physicochimical.txt", sep=";")
# Exporte table to work on it later
```
  
Then I analyse standard error by drawing.  
```{r plot standard error, echo=TRUE, message=FALSE, warning=FALSE, include=TRUE}
Z<-read.table("Physicochimical.txt", header = TRUE, sep = ";")
o <- order(Z$Date)
Z<-Z[o,]
Z$Site<-as.factor(Z$Site)

for (i in c(4,6,8,10))
{
  se<-ggplot(Z) + geom_point(aes(x = Date, y = Z[,i], color = Site)) + theme_minimal() + xlab("Date") + ylab(variable.names(Z)[i])
  print(se)
}
```
Now, we can see some standard error were higher than other (example for Pb one point has error higher than 0.1), so I will verify these specific data.  
This method allows me to found and correct five misprint in the data. I will take in mind other outlier to check their influences on the futur analyses.  

# Verify the repeatability of the data  
As in the previous [blog article](https://mvaugoyeau.netlify.com/post/my-first-article/), I verify the repeatability of the corrected data with simple linear model.  
```{r repeatability, echo=TRUE, message=FALSE, warning=FALSE}
Y$rep<-paste(as.factor(Y$Site), as.factor(Y$Date))

# The repeatability of the pH for exemple
mod<-lm(Y[,4]~Y$rep)
anova(mod)
```
  
Next time, I will describe how I do data exploration and first analyse.  
  
