---
title: Principal Component Analysis
author: "Marie Vaugoyeau"
date: '2018-06-13'
slug: principal-component-analysis
categories:
- stat
- graphique
tags:
- stat
- graph
featuredpath: "/images/featured/"
featured: "Fleur.jpg"
featuredalt: "Principal Component Analysis"
linktitle: "Principal Component Analysis"
images: 
  - /img/featured/Fleur.jpg  
description : "How I do Principal Component Analysis and choice of n factors" 
output:
  blogdown::html_page:
    toc: true
---
  
With [article about correlations](https://mvaugoyeau.netlify.com/post/start-with-the-data/), we saw data from airquality were correlated.  
Sometimes it is need to use Principal Component Analysis (PCA) to determine non correlated variables in order to analyze data.  
It is the subject of this blog article and especially, how many new variables were needed.  

# PCA  
As previously I use airquality as data.  
To do PCA, I use the package **FactoMineR**.  
```{r PCA, echo=TRUE, message=FALSE, warning=FALSE, include=TRUE}
library(FactoMineR)
D<-airquality

pca<-PCA(D)
pca$eig
```  
  
The question is how much dimensions do we need to keep?  
  
The wonderful package **psycho** of  [Dominique Makowski](https://neuropsychology.github.io/psycho.R/2018/05/24/n_factors.html) has the response. Thank him!  
Updated, the `n_factors()` function now belongs to the `parameters` package.  
  
# Number of factor retained by psycho::n_factors()  
  
```{r nfacteurs, echo=TRUE, message=FALSE, warning=FALSE, include=TRUE}
library(magrittr)
library(psycho)

choice <- D %>% parameters::n_factors()
choice
summary(choice)
plot(choice)
```  
  
On the plot which shows the summary, you can see in yellow, the number of methods. The red line is the Eigenvalues and the blue line, the cumulative proportion of explained variance.  
According to this method, we can keep the two first dimensions from PCA.  
  
# Extraction of the variables  
dimdesc from **FactoMineR** gives correlations and *p*-value.  
X is the new data comes from PCA.  
```{r extraction donnees, echo=TRUE, message=FALSE, warning=FALSE, include=TRUE}
dimdesc(pca, axes = 1:2)

X<-cbind(pca$ind$coord[,1], pca$ind$coord[,2]) %>% set_colnames(c("PC1", "PC2"))
head(X)
```
  
And you, how do you choice the number of factors kept from PCA?  
