---
title: Automate the opening of Excel files in R
author: "Marie Vaugoyeau"
date: '2018-10-10'
slug: automate-the-opening-of-excel-files-in-r
categories:
- tips and tricks
tags:
- tips and tricks
featuredpath: "/images/featured/"
featured: "Lego.jpg"
featuredalt: "Automate the opening of Excel files in R"
linktitle: "automate-the-opening-of-excel-files-in-r"
images: 
  - /img/featured/Lego.jpg  
description : "How to automate the reading and opening of Excel file (or csv or other) in R" 
output:
  blogdown::html_page:
    toc: true
---
  
After a long period without an article, I decided to explain how I automate the opening and reading of `.xlsx`, `.csv` or other files in R.  
If you have other methods or remarks, [do not hesitate to contact me](https://mvaugoyeau.netlify.com/contact/).  
  
# Creating dummy datasets  
I decided to create dummy Excel files and then import them.  
  
```{r fichiers_excel_factices, echo=TRUE, message=FALSE, warning=FALSE, include=TRUE}

openxlsx::write.xlsx(airquality, file = "airquality.xlsx")
openxlsx::write.xlsx(iris, file = "iris2.xlsx")
openxlsx::write.xlsx(Orange, file = "orange.xlsx")
openxlsx::write.xlsx(mtcars, file = "mtcars3.xlsx")

```
  
# Importing files  
Now I'm going to re-import these files, I'll also go further by using a regular expression (regex) to have only files that have a number at the end of their names.  
I am not a specialist in regular expressions. If you know more about this, I recommend the article writed in french by [Lise Vaudor](http://perso.ens-lyon.fr/lise.vaudor/strings-et-expressions-regulieres/) but I am sure there are a lot of very good articles in english!    
  
```{r lire_fichier_excel, echo=TRUE, include=TRUE, message=FALSE, warning=FALSE}
library(tidyverse)

fichiers_excel <-  list.files(path = here::here("content", "post"), full.names=TRUE) %>% 
  str_subset(".xlsx") %>% 
  map(openxlsx::read.xlsx) %>% 
  set_names(list.files(path = here::here("content", "post"), full.names=TRUE) %>%
              str_subset(".xlsx") %>% 
              str_extract("\\w+.xlsx") %>% 
              str_remove(".xlsx"))

str(fichiers_excel)
```
  
Now we only want Excel files whose name has a number at the end.  
  
```{r lire_fichier_excel_sans_majuscule, echo=TRUE, include=TRUE, message=FALSE, warning=FALSE}

fichiers_excel_bis <-  list.files(path = here::here("content", "post"), full.names=TRUE) %>% 
  str_subset("[0-9].xlsx") %>% 
  map(openxlsx::read.xlsx) %>% 
  set_names(list.files(path = here::here("content", "post"), full.names=TRUE) %>%
              str_subset("[0-9].xlsx") %>% 
              str_extract("\\w+.xlsx") %>% 
              str_remove("[0-9].xlsx"))

str(fichiers_excel_bis)
```
  
Well I admit it, it does not represent a crazy interest but it is just to expose you the principle.  
Of course, it's possible to do the same thing with `.csv` files using`read.csv ()` and putting ".csv" in regex.  
  
Then you can complicate it by searching for files with specific names like the place and date of sampling or the output of measuring devices.  
If we take the sampling file case with names of this form "Paris20-03-2014", "Foljuif10-04-2012", the regular expression would be "([A-z] +) \\ w + ([0 -9] {2} [: punct:] [0-9] {2} [: punct:] [0-9] {4})".    
  
Here, hoping that this article is a little useful!  
  
