---
title:  Automatic generation and dispatch of media to customers, blocking points and forwarding solutions  
author: "Marie Vaugoyeau"
date: '2020-07-22'
slug: presentation-off-women-international-day
categories:
- graphique
- stat
- tips and tricks
tags:
- graphique
- stat
- tips and tricks
featuredpath: "/images/featured/"
featured: "immeuble_ny.png"
featuredalt: "Creating custom Power Point media with {officeR} and automatic emailing to clients with {sendmailR}"
linktitle: "presentation-off-women-international-day"
images: 
  - /img/featured/immeuble_ny.png  
description : "Article from the presentation that should have been made for the International Women's Day in Nantes on March 19 and which was cancelled due to the COVID-19 outbreak" 
output:
  blogdown::html_page:
    toc: true
---
  
```{r setup, include=FALSE}

knitr::opts_chunk$set(warning = FALSE, message = FALSE)

```

  
# Aim  
  
The aim is to create a communication support for clients to see the evolution of their results compared to the data of other clients in the same category (`Benchmark`).  
This support had to be generated and sent automatically to the customers.  
  
So that people who wish to reuse these lines of code, the template used is [here](/post/template_rstudio_article_off_eng.pptx), the data are those of the Penguins of the package `FlexParamCurve` by [Stephen OSWALD](https://www.rdocumentation.org/packages/FlexParamCurve/versions/1.5-5/topics/penguin.data) and [LifeCycleSavings](https://www.rdocumentation.org/packages/datasets/versions/3.6.2/topics/LifeCycleSavings) of the package `datasets`. Examples of the published media at the end of the article are available [here](/post/presentation_0993421052631561 308_eng.pptx) and [here](/post/presentation_0993421052631561 422_eng.pptx).  
  
Work sequence:  

_ [Creation of a template under Power Point](/post/template_rstudio_article_off_eng.pptx)   
_ Creation of the graphics to be inserted  
_ Creation `Power Point` support using the `officer` package from [David Gohel](davidgohel.github.io/officer/)  
_ Sending support with `sendmailR` from [Olaf Mersmann](www.rdocumentation.org/packages/sendmailR)  
  
  
# Creating graphs on the data  
  
When I carried out this work it was on customer data and therefore confidential. For this blog post I'm going to use the mass data of penguins according to their ages and compare it to the average data of other penguins born in the same order and the same year to show how to make a graph with a benchmark. I'm also going to do another graph on the international data of real disposable income per capita.  
  
## Comparison of one penguin to another  
  
Comparing a penguin to others of the same year and birth order is easy with `ggplot`.  
  
```{r realisation_d_un_graphique}

library(tidyverse)
library(FlexParamCurve) # to get the data on the penguins

creation_of_the_graph_for_comparing_a_penguin_to_others <- function(num_penguin){
  
  benchmarck <- penguin.data %>% 
    filter(
      bandid != num_penguin,
      year == (penguin.data %>% filter(bandid == num_penguin) %>% slice(1))$year,
      ck == (penguin.data %>% filter(bandid == num_penguin) %>% slice(1))$ck
    ) # Benchmarking is done with data from other penguins born in the same year and in the same hatching order
  
  penguin.data %>% 
    filter(
      bandid == num_penguin
    ) %>% 
    ggplot() +
    aes(
      x = ckage,
      y = weight
    ) +
    geom_smooth(data = benchmarck, color = "red", fill = "orange", alpha = 0.4, formula = y ~ x^2) + 
    geom_point(color = "black") + 
    geom_smooth(se = FALSE, formula = y ~ x^2, color = "black") +
    theme_classic() +
    ggtitle(glue::glue("Penguin '{num_penguin}' compare to his peers (in red)"))
  
 }

"-0.993421052631561 422" %>% 
  creation_of_the_graph_for_comparing_a_penguin_to_others()

"-0.993421052631561 308" %>% 
  creation_of_the_graph_for_comparing_a_penguin_to_others()

```
  
  
  
## Creating a function to view data in "French" format  
  
Basically, R presents numerical data without spaces or rounding but from time to time in graphs, it may be interesting to be able to display the figures in "French" format i.e. `123 456 789` as Excel can do natively.  
  
For this example, I will use the data from LifeCycleSavings.  
  
```{r mise_en_forme_française}

# library(tidyverse) # not useful if already called previously
library(scales) # to use the label_number function for number formatting
library(ggrepel) # to add non-overlapping values to the chart

fr_formatting <- label_number(accurancy = 1)
# I ask that there be no decimal point

# it works 
1332859475.645878 %>% fr_formatting()

LifeCycleSavings %>% 
  rownames_to_column("country") %>% 
  mutate(label = dpi %>% fr_formatting()) %>% 
  ggplot() +
  aes(x = country, y = dpi) +
  geom_segment(aes(xend = country, y = 0, yend = dpi), size = 6) +
  geom_label_repel(aes(label = label), segment.color = NA, direction="y", size = 3) +
  theme_classic() +
  theme(
    axis.text.x = element_text(angle = 90)
  )
# ouch the rounding is not done

```
  
You can see that despite my preliminary test, it doesn't work well...  
Indeed, if the global formatting is good, the rounding is not done... But why????  
  
I was stuck for a while wondering what was wrong and why and ended up asking for help on the French slack of R users, [GRRR](https://lstu.fr/rgrrr).  
**Little info point**: Whatever media you want to ask for help on, think about making a `repex`, a reproducible example!  
Yes, people on the forums or slacks are your friends and yes, the R sphere is benevolent but don't forget that nobody is in your head (except you, at least normally), so the first step towards solving a problem is to take it out of context so that you can explain it to everyone using accessible data!  
This has already happened to me more than once that when I did my repex to ask for help I saw where the problem was.   
  
Here are the lines of code I shared on the slack.  
  
```{r creation_exemple_reproductible}

# library(scales)
# library(tidyverse)

fr_formatting <- label_number(accurancy = 1)

c(13556.4646, 546946.65465) %>% map(fr_formatting)

tibble(
  valeur = 
    runif(n = 50, min = 100, max = 6000)
  ) %>% 
  mutate(mise_en_forme = valeur %>% fr_formatting)

```
  
This time it was [Julien Barnier](https://twitter.com/lapply) who helped me out by pointing out that I had made a mistake at `accuracy`, there is no 'n'....   
Strangely enough, it worked better right away.     
  
```{r reponse_au_probleme}
fr_formatting <- label_number(accuracy = 1)

tibble(
  valeur = 
    runif(n = 50, min = 100, max = 6000)
  ) %>% 
  mutate(mise_en_forme = valeur %>% fr_formatting)

graphique_LCS <- LifeCycleSavings %>% 
  rownames_to_column("country") %>% 
  mutate(label = dpi %>% fr_formatting()) %>% 
  ggplot() +
  aes(x = country, y = dpi) +
  geom_segment(aes(xend = country, y = 0, yend = dpi), size = 6) +
  geom_label_repel(aes(label = label), segment.color = NA, direction="y", size = 3) +
  theme_classic() +
  theme(
    axis.text.x = element_text(angle = 90)
  )

graphique_LCS

```
  
  
  
# Creating PowerPoint support with `officer`  
  
Basically I'm not too much for creating `pptx` support with R, I prefer `html` reports but it is still a much awaited support for communication to customers.  
  
With the [`officer`](davidgohel.github.io/officer/) package, I will create simple but dynamic support with links between slides.  
  
  
```{r creation_du_support}

library(officer)
library(glue)

location_of_support <- "template_rstudio_article_off_eng.pptx"

# pour voir les types de diapos qui existent et les éléments qui les constituent
read_pptx(location_of_support) %>%
  layout_properties() %>% 
  select(master_name, name, type, id, ph_label)

creating_support <- function(num_penguin){
  read_pptx(location_of_support) %>% 
    # 1. Creating the title slide
    add_slide(
      layout = "Diapositive de titre", 
      master = "Intégral"
    ) %>% 
    ph_with(
      value = num_penguin, 
      location = ph_location_label("Title 1")
    ) %>% 
    ph_with(
      value = penguin.data %>% filter(bandid == num_penguin) %>% slice(1) %>% as_tibble() %>% select(year, ck),
      location = ph_location_label("Espace réservé du tableau 4")
    ) %>% 
    
    # 2. Adding clickable summary
    add_slide(
      layout = "Resume",
      master = "Intégral"
    ) %>% 
    ph_with(
      value = glue("Penguin identification number : {num_penguin}"),
      location = ph_location_label("Espace réservé du texte 5")
    ) %>% 
    ph_with(
      value = "Information about Marie Vaugoyeau",
      location = ph_location_label("Espace réservé du texte 9")
    ) %>% 
    ph_with(
      value = "Graphique Life Cycle Saving",
      location = ph_location_label("Espace réservé du texte 11")
    ) %>% 
    ph_with(
      value = glue("Graph associated with the penguin {num_penguin}"),
      location = ph_location_label("Espace réservé du texte 13")
    ) %>% 
    ph_with(
      value = "Contact information",
      location = ph_location_label("Espace réservé du texte 15")
    ) %>% 
    
    # 3. Addition of the information slide about me (small ad page)
    add_slide(
      layout = "PrésentationMarie",
      master = "Intégral"
    ) %>% 
    ph_with(
      value = "Click back to the menu", 
      location = ph_location_label("Espace réservé du contenu 5")
    ) %>%
    ph_slidelink(
      ph_label = "Espace réservé du contenu 5", 
      slide_index = 2
    ) %>% 
    
    # 4. Adding LCS graph
    add_slide(
      layout = "Titre et contenu",
      master = "Intégral"
    ) %>% 
    ph_with(
      value = "Graph LCS",
      ph_location_label("Title 1")
    ) %>% 
    ph_with(
      value = graphique_LCS,
      ph_location_label("Content Placeholder 2")
    ) %>% 
    ph_with(
      value = "Click back to the menu", 
      location = ph_location_label("Espace réservé du contenu 5")
    ) %>%
    ph_slidelink(
      ph_label = "Espace réservé du contenu 5", 
      slide_index = 2
    ) %>%
    
    # 5. Adding custom penguin graphic
    add_slide(
      layout = "Titre, Contenu et Image",
      master = "Intégral"
    ) %>% 
    ph_with(
      value = "Penguin graph",
      ph_location_label("Title 1")
    ) %>% 
    ph_with(
      value = external_img("pingouin.png"), # image extraite de l'article à l'origine des données
      location = ph_location_label("Espace réservé du contenu 3")
    ) %>% 
    ph_with(
      value = glue("{(penguin.data %>% filter(bandid == num_penguin) %>% slice(1))$year} {(penguin.data %>% filter(bandid == num_penguin) %>% slice(1))$ck}"),
      ph_location_label("Espace réservé du contenu 8")
    ) %>% 
    ph_with(
      value = creation_of_the_graph_for_comparing_a_penguin_to_others(num_penguin),
      ph_location_label("Espace réservé du contenu 6")
    ) %>% 
    ph_with(
      value = "Click back to the menu", 
      location = ph_location_label(ph_label = "Espace réservé du contenu 5")
    ) %>%
    ph_slidelink(
      ph_label = "Espace réservé du contenu 5", 
      slide_index = 2
    ) %>%
    
    # 6. Adding the contact slide
    add_slide(
      layout = "Contact",
      master = "Intégral"
    ) %>% 
    ph_with(
      value = "Click back to the menu", 
      location = ph_location_label(ph_label = "Espace réservé du contenu 5")
    ) %>%
    ph_slidelink(
      ph_label = "Espace réservé du contenu 5", 
      slide_index = 2
    ) %>%
    
    # Creating links on the contents page
    on_slide(index = 2) %>%
    ph_slidelink(
      ph_label = "Espace réservé du texte 9", 
      slide_index = 3
    ) %>%
    ph_slidelink(
      ph_label = "Espace réservé du texte 11", 
      slide_index = 4
    ) %>%
    ph_slidelink(
      ph_label = "Espace réservé du texte 13", 
      slide_index = 5
    ) %>%
    ph_slidelink(
      ph_label = "Espace réservé du texte 15", 
      slide_index = 6
    )
}

```
  
  
# Sending the support with the package `sendmailR`  
  
Once the skeleton of the support is ready, all that's left to do is to send each penguin (or each client) his personalized support.  
  
```{r envoi_du_support, eval = FALSE}

library(sendmailR)

num_penguin <- c("-0.993421052631561 422", "-0.993421052631561 308")

message_support <- map(
  .x = num_penguin,
  .f = ~ mime_part(
    creating_support(.x), 
    "support.pptx"
  )
)

message_support[[1]]$`Content-Type` <- "application/vnd.openxmlformats-officedocument.presentationml.presentation"

message_support[[2]]$`Content-Type` <- "application/vnd.openxmlformats-officedocument.presentationml.presentation"

shipping_address <- c("experimentatateur1@truc.com", "experimentateur2@truc.com") # address of the person who followed the penguin/customer

penguin_address <- c("pingouin1@pole_nord.fr", "pingouin2@pole_nord.fr") # email address of the penguin/customer

# Sending emails
map2(
  .x = shipping_address,
  .y = penguin_address,
  .f = ~ sendmail(
    from = glue("<{.x}>"),
    to = glue("<{.y}>"),
    subject = glue("Synthesis of penguin weight gain {num_penguin}"),
    msg = list(
      glue("Hello,\nPlease find attached the synthesis for the penguin {num_penguin}.\nHave a nice day, \nMarie Vaugoyeau"),
      message_support
    ),
    control = list(smtpServer = "ASPMX.L.GOOGLE.COM")
  )
)

```
  
Oops, it doesn't work...  
But why??   
  
In fact, a quick phone call to [David Gohel](https://twitter.com/davidgohel), 1 minute of conversation and hop the solution is obvious, what I treat as a pptx is in fact an rpptx....  
So you have to edit the pptx support before sending it and then it's ok !  
  
Two solutions, either we save all the pptx supports we send for archive (solution taken [here](/post/presentation_0993421052631561 308_eng.pptx) and [here](/post/presentation_0993421052631561 422_eng.pptx)), or we create a temporary support which will be overwritten.    
   
```{r edition_du_support, results = FALSE}

pptx_support_recording <- function(num_penguin){
  num_penguin %>% 
    creating_support() %>% 
    print(glue('presentation_{num_penguin %>% str_remove_all("[:punct:]")}_eng.pptx'))
}

num_penguin <- c("-0.993421052631561 422", "-0.993421052631561 308")

num_penguin %>% 
  map(pptx_support_recording)

```
  
  
There you go!  
  
I hope that this article will help you better seek help and make your work with R easier.   
  
  
